---
title: "dataProcessing"
author: "Thao Phuong Pham"
date: "1/5/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```

**Bảy Núi An Giang Project**

Raw data -> **data processing** -> data analyzing -> communicating results


The data processing contains these works:  

1. Load the raw dataset  
2. Display structure and overall information of the data set: using `str()`  
3. From the result of `str()`, decide which variables need to be modified  
4. Dealing with missing values  
5. Using knowledge to check for suspected wrong data  
6. Tidy data  
7. Subset data as needed  
8. Write data to external files. 

**Library**

```{r load-library, warning=FALSE, message=FALSE}
library(dplyr)
library(data.table)
library(stringi)
library(tidyselect)

```


## 1. Load the raw dataset

Using MS Excel to look at the data interactively and there are things that I need to take care of:  

1. A title (row 1) that need to be remove  
2. Second row containing header (columns' name) that are not suitable for data analyzing

Using `readxl::read_xlsx()` to load the data.

```{r load-dataset}
AnGiangData <- readxl::read_xlsx("data/rawAnGiangData.xlsx", sheet = 1, skip = 2, col_names = FALSE)

```

Create a `variable.txt` file containing all the names of the dataset's columns, load it to R and use `colnames()` to assign names.
```{r rename-column}
variable <- read.table("data/variable.txt")
colnames(AnGiangData) <- variable$V1
```

## 2. Display structure and overall information of the data set: using `str()`
```{r data-structure}
str(AnGiangData)

```

`str()` gives result that:

- the data type of some variables are wrongly decided.  
- the dataset contains non-English text that cannot be presented in R.  
- text contains a mix of uppercase and lowercase.  

## 3. From the result of `str()`, decide which variables need to be modified 

### uppercase and convert all the text to Latin

At first, I decided to work with the objects' data type and then moving to manipulate text case and coding style.
However, I quickly found out that, `stri_trans_general()` and `toupper` convert all things to character type.
So it's better to work with the data type at the end.
```{r text-cleaning}
AnGiangData <- AnGiangData %>% 
  mutate(across(everything(), stri_trans_general, "latin-ascii")) %>% 
  mutate(across(everything(), toupper))

```

### decide data types for the objects

I manually create text file (outside R) called `dataType.txt` that contains the right data type for all the variables.
Read the data into R and use it to correct all objects that have wrong data type.

```{r data-type-modify}
dataType <- read.table("data/dataType.txt", header = F)

charIndex <- which(dataType$V1 == "character")
AnGiangData[,charIndex] <- lapply(AnGiangData[,charIndex], as.character)

numIndex <- which(dataType$V1 == "numeric")
AnGiangData[,numIndex] <- lapply(AnGiangData[,numIndex], as.numeric)

facIndex <- which(dataType$V1 == "factor")
AnGiangData[,facIndex] <- lapply(AnGiangData[,facIndex], factor)

```

### Remove unused rows and columns

There are two duplicated columns (`income.2018.1` and `income.2018.2`). Remove one.

There are 3 rows at the end of the dataset that is NULL.

```{r remove-row-column}
AnGiangData <- AnGiangData %>% 
  select(!c(income.2018y)) %>% 
  slice(1:181)

```




